### 2023-09-19 manuscript/wp4.tex
ms 19/09

\PassOptionsToPackage{hidelinks}{hyperref}
\documentclass[doc,biblatex]{apa7}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{references.bib}
\captionsetup[figure]{textfont=rm,font=small,labelfont=bf}
\usepackage{amsmath}
\usepackage{mathspec}
\setallmainfonts{Times New Roman}
\setmainfont{Times New Roman}

\title{Why do languages tolerate heterography? An experimental investigation of the cultural evolution of informative orthography}

\shorttitle{The cultural evolution of informative orthography}

\authorsnames[1,1]{Jon W. Carr, Kathleen Rastle}

\authorsaffiliations{{Department of Psychology, Royal Holloway, University of London, England}}

% \authornote{\textbf{Cite as:} Carr, J.\@ W., \& Rastle, K.\@ (2023). Informativity in the writing system. \textit{OSF Preprints}. Version~0.}

\abstract{It is widely acknowledged that opaque orthographies place additional demands on learning, often requiring many years to acquire fully. It is less widely recognized, however, that such orthographies may offer certain benefits in the context of reading. For example, the heterographic homophones (e.g., <knight> vs. <night>) impose additional costs in learning but reduce ambiguity in reading. Could it be the case, then, that heterography---and therefore orthographic opacity---may sometimes be selected for in the evolution of writing systems, despite the cost in learning? We investigate this question by simulating the evolution of orthography as it is transmitted from one generation to the next, both with and without a communicative pressure for avoiding ambiguity. In addition, we consider two mechanisms by which informative heterography might be selected for: \textit{differentiation}, in which new spellings are created to differentiate words that are homophonous in speech (e.g., <lite> from <light> to signal low-fat), and \textit{conservation}, in which heterography arises through exaptation as an epiphenomenon of sound change (e.g., <meat> vs. <meet> resulting from the /ɛː/–/eː/ merger in the Great Vowel Shift). Under pressure from learning alone, orthographic systems become transparent, but when combined with communicative pressure, these systems can become informative, diverging from the spoken language in order to express meaning directly.}

\keywords{communication; cultural evolution; informativity; iterated learning; language evolution; orthography; spelling; sound change; writing}

\begin{document}

\maketitle

\noindent
Written languages represent spoken languages. However, there are many ways in which written languages diverge. E.g. spacing between words helps readers parse a sentence even if they are not technically required. Written languages can also have their own regularities such as spelling related words similarly (magician), past tense spellings directly conveying meaning, certain spellings conveying word class ous. These features abound in writing, especially in languages such as English that have not been subject to serious regulation.

One notable case of this is the heterographic homophones---words that sound alike but which are written with different spellings (e.g., <meat> and <meet>, both pronounced /miːt/). Heterographic spellings such as there can be useful in reading. For example, faced with a sentence beginning /ðɛr/, a listener will have high uncertainty about what word---or even what sentence structure---is likely to come next: a noun, as in /ðɛr dɒg/, a form of the verb to be, as in /ðɛr ɪz/, or the progressive form of a verb, as in /ðɛr gəʊɪŋ/. In writing, by contrast, this uncertainty is greatly reduced; the spellings <their>, <there>, and <they're> differentiate these cases, giving the reader a headstart on parsing the upcoming syntax. Heterography may, therefore, confer certain benefits in reading, particularly in terms of comprehension. Heterography permits a writing system to be---at least in some areas---more informative than its spoken counterpart, avoiding ambiguity.

Irish orthography has special features to handle the initial consonant mutations that occur extensively across the Celtic languages. For example, under certain grammatical contexts, a word such as /pɑʃti/ (children) will be transformed into /bɑʃti/, a phenomenon known as eclipsis. The writing system, however, does not simply replace the <p> with a <b>, as one might predict; instead, both letters are included, allowing the orthography to, at~once, show meaning and phonetic form: the eclipsed spelling of <páistí> is <bpáistí>. Moreover, in some cases, the eclipsed spellings result in ambiguity avoidance---in this case, if the <p> were not retained, the hypothetical spelling <báistí> would be confusable with the genitive form of the word for rain.

Even the most orthographically transparent of languages have at least some techniques for disambiguating homophones in writing. Italian, for example, uses the grave accent to distinguish common homophonous words: <la> (\textit{the}) vs.\@ <là> (\textit{there}). Many other languages employ various diacritics for the same purpose, such as the acute accent in Danish---<en> (\textit{a}) vs.\@ <én> (\textit{one})---or the circumflex in Serbo-Croatian---<da> (\textit{yes}) vs.\@ <dâ> (\textit{gives}).

However, this all sounds great but additional informativeness comes at a cost. A system of words that conveys more information will---in general---be more complex than one that conveys less information. There exists a tradeoff between the simplicity of the spelling system and its ability to be expressive. A simple spelling system is one that is easy to learn, use, and process; for example, by being transparent with respect to phonology. The idea of a tradeoff between simplicity and informativeness in language structure has a longstanding history \parencite{Gabelentz:1891, Zipf:1949, Martinet:1952, Rosch:1978} and has recently been explored in a wide range of both typological \parencite{KempRegier:2012, Kemp:2018} and experimental \parencite{Kirby:2015} studies. Of particular note here is the idea that the pressure for simplicity in language structure derives principally from learning, while the pressure for informativeness derives from the need for precise communication.

Unpack some of this literature a bit.

Any explanation for heterographic homophones must, therefore, demonstrate that heterographic spellings convey a benefit in terms of informativeness that is outweighed by the cost of learning the (often arbitrary) spelling distinction. What are the benefits? There seem to be broadly two kinds of benefit. (1) disambiguation in the writing of words that would otherwise be ambiguous if spelled transparently. (2) permitting the cognitive system to rapidly and accurately attain the meaning of a word. In this paper we limit ourselves to informativeness of the first kind, which is more amenable to experimental testing.

Is it possible, then, that heterography may be selected for in the cultural evolution of orthography? And if so, what might the mechanism be through which such selection could operate? \textcite[pp.~325--326]{Berg:2021} sketch two models of how a word might enter a state of heterography, as depicted in Fig.~\ref{models_of_heterography}. The first model, \textit{the differentiation model}, explains heterography through the creation of a new orthographic forms. For example, the word \textit{lite} is a respelling of \textit{light} frequently used in food products to distinguish light-in-calorific-weight from other meanings of \textit{light}; in British English, the word \textit{cheque} (perhaps influenced by \textit{exchequer}) differentiates the bank draft from other meanings of the word \textit{check}; and the word \textit{byte} was a deliberate respelling of \textit{bite} to avoid accidental mutation into the closely related term \textit{bit} \parencite{Buchholz:1977}. Etymological (and folk-etymological) spellings introduced during the Renaissance also occasionally resulted in heterographic homophones, perhaps contributing to their success: \textit{scene}--\textit{seen}, \textit{scent}--\textit{sent}, \textit{whole}--\textit{hole} \parencite[pp.~58--59]{Scragg:1974}, and many monosyllabic words that are homophonous with common function words have tended to adopt alternate spellings: \textit{be}--\textit{bee}, \textit{but}--\textit{butt}, \textit{by}--\textit{bye}--\textit{buy}, \textit{for}--\textit{fore}--\textit{four}, \textit{in}--\textit{inn}, \textit{or}--\textit{oar}--\textit{ore}, \textit{so}--\textit{sew}, \textit{to}--\textit{too}--\textit{two}, etc. Heterography is also very common in proper nouns: Surnames that are homophonous with common nouns often add <e> (\textit{Clarke}, \textit{Greene}, \textit{Wilde}) or double the final consonant (\textit{Carr}, \textit{Hogg}, \textit{Mann}), and it is common for brand names to adopt creative spellings, such as \textit{flickr}, \textit{Froot Loops}, or \textit{MVMT}.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/models_of_heterography.eps}}
	\vspace*{2pt}
	\caption{Two models of heterography. In the differentiation model, two meanings (represented by pink and blue) are, at time $T_1$, expressed by a single phonetic form $P_1$ and a single orthographic form $O_1$; however, by time $T_2$, two orthographic forms have emerged to differentiate the meanings in writing. In the conservation model, the two distinct phonetic forms that existed at time $T_1$ have become homophonous by time $T_2$, but the two corresponding orthographic forms have been conserved, resulting in the same state of heterography as in the differentiation model. Adapted from \textcite[pp.~325--326]{Berg:2021} with permission.}
	\label{models_of_heterography}
	\end{figure}

The second model, \textit{the conservation model}, explains heterography as the historical residue of sound change: Two (or more) spoken forms merge and become homophonous, but the original heterographic spellings are conserved in the orthography. In English, for example, the meat--meet merger ultimately resulted in Middle English /ɛ:/ (spelled <ea>) and /e:/ (spelled <ee>) merging into /i:/ in Early Modern English, although the spellings were never updated accordingly, thus giving rise to a set heterographic homophones that persist in present-day English: \textit{heal}--\textit{heel}, \textit{leak}--\textit{leek}, \textit{meat}--\textit{meet}, \textit{read}--\textit{reed}, \textit{sea}--\textit{see}, \textit{team}--\textit{teem}, \textit{weak}--\textit{week}, etc. Similarly, other historic sound changes have likely resulted in (or contributed to) pairs of words entering a state of heterography, such as the reduction of /kn/ into /n/ (\textit{knight}--\textit{night}, \textit{know}--\textit{no}, \textit{knot}--\textit{not}, etc.), the loss of /ç/ (\textit{eight}--\textit{ate}, \textit{right}--\textit{rite}, \textit{sight}--\textit{site}, etc.), and the merger of /ʍ/ into /w/ (\textit{whale}--\textit{wail}, \textit{which}--\textit{witch}, \textit{whine}--\textit{wine}, etc.).

In this paper, we simulate the processes of differentiation and conservation experimentally using the experimental iterated learning paradigm \parencite{Kirby:2008}. In this paradigm, some linguistic system is passed along a \textit{transmission chain} of human participants, resulting in repeated cycles of learning and production. Participant $i$ in a transmission chain learns the system based on the linguistic output of participant $i-1$ and then produces new linguistic output for participant $i+1$ to learn from. Following several \textit{generations} of this process, the linguistic system can gradually adapt to the learning biases of the human learners, yielding emergent linguistic phenomena, such as compositionality \parencite{Kirby:2008, Kirby:2015, Beckner:2017}, combinatoriality \parencite{Verhoef:2015}, category structure \parencite{Carr:2017, Canini:2014, Silvey:2019}, and regularization \parencite{Smith:2010, Ferdinand:2019}. For reviews, see \textcite{Tamariz:2017, Kirby:2014, ScottPhillips:2010}.

Bottleneck

Preregistration: \url{https://aspredicted.org/YLG_CKD}

%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%

\section{Experiment 1}

Our first experiment tests the ability of the differentiation model to explain the emergence of informativeness in orthography. In particular we had two main hypotheses:

\subsection{Methods}

To simulate the cultural evolution of orthography, we adopt the experimental iterated learning paradigm \parencite{Kirby:2008, Kirby:2015}. In this paradigm, some linguistic system is passed along a \textit{transmission chain} of human participants, resulting in repeated cycles of learning and production. Participant $i$ in a transmission chain learns the system based on the linguistic output of participant $i-1$ and then produces new linguistic output for participant $i+1$ to learn from. To explore the hypotheses outlined above, we conducted the experiment under two conditions: Transmission-only and Transmission~+ Communication.

[make this more specific to the present experiment - what is passed down the chain - preview the methods coming up]

\subsubsection{Participants}

We recruited 287 participants via the Prolific platform. Participants were paid £2.00 for participation plus additional bonuses of up to £1.08 as detailed below (median bonus: £0.74). The median completion time was 15~m with a median hourly rate of £8.05 (£10.94 including bonus). We limited recruitment to (self-declared) native English speakers, since it was important that participants would perceive the spoken forms in a relatively consistent way (particularly in the case of Experiment~2). 14 participants were excluded because they (or their communication partners) used English color words (8) or failed the auditory attention checks (6). A further three participants were lost to communication-game pairing failures. The final dataset comprises 270 participants: 90 in the Transmission-only condition (10 chains of 9 participants) and 180 in the Transmission~+ Communication condition (10 chains of 9 pairs of participants).

\subsubsection{Stimuli}

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/stimuli.eps}}
	\vspace*{2pt}
	\caption{The nine alien object stimuli with their stems and suffixes. The spoken and written forms of the stems were fixed and unchanging throughout the experiment, as were the spoken forms of the suffixes, which were always homophonous; however, the written forms of the suffixes were free to evolve over time, potentially taking on differentiated forms to indicate color (e.g., <-co>, <-ko>, and <-qo> to represent pink, yellow, and blue).}
	\label{stimuli}
	\end{figure}

Participants were taught words for nine ``alien'' objects---three shapes (pentagon, star, torus) in three colors (pink, yellow, blue), as depicted in Fig.~\ref{stimuli}. The alien words had a spoken and written form composed of a stem and suffix. The stems, which always express the shape dimension, were /buvɪ/ <buvi> (the pentagon), /zɛtɪ/ <zeti> (the star), and /wɒpɪ/ <wopi> (the torus). These stems were fixed and unchanging throughout both experiments reported in this paper and were designed to be easy to learn by being graphically and phonetically iconic of the shapes they represent. Throughout Experiment~1, the spoken form of the suffix was always pronounced /kəʊ/, but its spelling was free to change over time. Thus, the spoken form of the language consists of just three unique words---/buvɪkəʊ/, /zɛtɪkəʊ/, and /wɒpɪkəʊ/---that mark only a shape distinction; however, the \textit{spelling} of the suffix could potentially take on different forms to mark color.

Each of the 10 transmission chains was seeded with a randomly-generated suffix spelling system, which was created by randomly mapping the following nine spellings onto the nine objects: <co>, <coe>, <coh>, <ko>, <koe>, <koh>, <qo>, <qoe>, <qoh>. In other words, the /k/ sound may be spelled <c>, <k>, or <q> and the /əʊ/ sound may be spelled <o>, <oe>, or <oh>, although the initial seed system contains no particular regularity. This procedure models an initial state of high spelling variation (every object has a unique suffix spelling), but these spellings may, over time, become transparent (the /k/ and/or /əʊ/ sounds take on consistent spellings) or expressive (spellings of /k/ and/or /əʊ/ become systematically associated with meaning). The spoken forms were synthesized using the Apple text-to-speech synthesizer (Tessa voice).

\subsubsection{Transmission procedure}

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/transmission.eps}}
	\vspace*{2pt}
	\caption{(A) Transmission-only procedure. Each generation consists of a single participant, who first receives training on six of the nine suffixes and then produces suffixes for all nine items. These productions are then used as the training material for the next generation in the chain. The initial system of suffixes is randomly generated with high spelling variation; but by the ninth generation, the system is expected to become transparent. (B) Transmission~+ Communication procedure. Each generation now consists of two participants who engage in a communicative task. Both participants receive training on the same system from the previous generation. Under a communicative pressure, the suffix spellings are expected to become expressive of color, despite the spoken forms being homophonous.}
	\label{transmission}
	\end{figure}

Participants were arranged into transmission chains such that the spellings produced by one participant would subsequently be taught to the next participant in the chain (see Fig.~\ref{transmission}A). The first participant in a chain was taught the initial, randomly generated seed system, and this system was then free to evolve as it was transmitted to subsequent generations. Importantly, this process was subject to a bottleneck on transmission: Not all nine spellings were transmitted from one generation to the next; rather, the participant at generation $i$ would observe only six of the nine spellings produced at generation $i-1$ (at least one of each shape and at least one of each color). Nevertheless, participants were asked to produce a spelling for all nine objects, meaning that generalization was required for three unseen items. Transmission continued for nine generations in each of ten independent chains. In the Transmission~+ Communication condition (see Fig.~\ref{transmission}B), each generation consists of a pair of participants, but the productions of only one of the participants (the primary participant; determined by whichever participant started the experiment first) are iterated to the next generation (the productions of the secondary participant are thus a cultural deadend). The role of the secondary participant is to act as a communicative partner for the primary participant, inducing pressure for the language to become informative.

\subsubsection{Training procedure}

All participants were trained on the spoken and written forms through a combination of passive exposure trials and ``mini-test'' trials. Participants were also told explicitly that the stems were iconic of the objects' shapes, allowing participants to focus on learning the suffixes during the training phase. In passive exposure trials, the alien objects were presented alongside the written and spoken forms in quick succession for 2~s each. In mini-test trials, which were interleaved among the passive exposure trials, the participant was asked to type the appropriate written form for an object and was given feedback on any errors (deleted characters shown in red strikethrough text and additions shown in bold green text). The participant received a 2p bonus for spelling the word correctly but had to submit their response within 20~s. Each of the six object--word pairs in the training set (i.e., the seen items that passed through the bottleneck on transmission) was passively exposed 18 times and mini-tested six times, resulting in a total of 108 passive exposure trials and 36 mini-test trials (the maximum bonus in training was therefore 72p). To check that participants were listening to the spoken forms, they were asked auditorily to click on the alien object at three random points during training; participants who did not follow this instruction were excluded.

\subsubsection{Test procedure in Transmission-only}

After training, participants assigned to the Transmission-only condition completed a test phase, alternating between production and comprehension trials. In production trials, the participant was shown an object and heard its associated pronunciation. The participant's task was to type the appropriate spelling. The input box was limited to eight lowercase Latin characters, and participants had to spell the stem correctly to continue to the next trial. Since participants heard the word pronounced aloud, typing the stem correctly should have been trivial, but in cases where the stem was initially spelled incorrectly, a popup message explicitly reminded the participant of the correct spelling of the stem. This restriction was imposed to prevent the stems from diverging from their spoken forms over time; however, no such restriction was imposed on the spelling of the suffix. Since the overall word length was restricted to eight characters and since all stems were four letters long, participants could use, at a minimum, a zero suffix and, at a maximum, a four-letter suffix. In comprehension trials, the participant was shown a word and had to click on the matching object from an array of all nine objects arranged in random order (in cases where multiple objects were described by the same wordform, any of the objects was a valid choice). In both types of test trial, the participant was awarded a bonus of 2p for each correct answer, but no explicit feedback was provided. Each of the nine object--word pairs (i.e., including unseen items) was tested once in production and once in comprehension, resulting in 18 trials (the maximum bonus in test was therefore 36p).

\subsubsection{Test procedure in Transmission~+ Communication}

Participants assigned to the communicative condition completed a live, over-the-internet communication game with another participant, who had received training on the same orthographic system inherited from the previous generation (albeit subject to a separate application of the bottleneck on transmission; i.e., the seen items were selected independently for each participant). The game closely mirrored the overall structure of the test administered to participants in the non-communicative condition, with the production and comprehension trials becoming the two sides of a single communicative interaction. On a given trial, one participant (the director) would complete a production trial under the same input restrictions described above (i.e., produce a form for a target meaning), and the word they used was relayed to the other participant (the matcher), who would then complete a comprehension trial in response to that word (i.e., pick an object from the matcher array). The two participants then switched roles, resulting in the same overall trial structure as the Transmission-only condition (i.e., alternation between production and comprehension trials and the same bonusing scheme).

The framing and goal of the communication game was, however, quite different from the non-communicative test. In communication, the shared goal of the director and matcher is to have a successful communicative interaction, not necessarily to reproduce what they had learned in training. Both participants received the 2p bonus each time an interaction was successful; that is, the reward structure is not based on using the ``correct'' forms taught in training but based on accurately conveying meaning. The director is thus incentivized to produce a wordform that is unambiguous, and the matcher is incentivized to carefully interpret what the director has attempted to convey. The second important difference from the non-communicative test was that participants received rich feedback on the interaction: The director saw which object the matcher clicked on, and the matcher saw which object was the correct target. Overall, then, the communication game is identical to the non-communicative test in terms of the task to be performed (nine productions and nine comprehensions), but the goal is quite different: In the non-communicative test, the goal and reward structure is based on accurately reproducing the orthography learned during training, whereas in the communication game, the goal and reward structure is based on successfully communicating a target object.

\subsection{Results}

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/dif_lrn.eps}}
	\vspace*{2pt}
	\caption{Results from the Transmission-only condition in Experiment~1 (differentiation). Each matrix shows the suffix spelling system in use at a particular generation (shape on the rows, color on the columns, as in Fig.~\ref{stimuli}). Chains are labeled A--J and generations are labeled 0–9 (0 is the randomly generated seed system). Each chain uses an independent color palette, with each color representing a particular suffix spelling; similar colors indicate similar spellings. Spellings in bold-italic are the generalizations on unseen items. By the ninth generation, most systems are degenerate (e.g., Chains~A, B, and~I), redundant (e.g., Chains~D, E, and~H), or a mixture of the two (e.g., Chain~F).}
	\label{dif_lrn}
	\end{figure}

The results from all ten chains (labeled A--J) in the Transmission-only condition are shown in Fig.~\ref{dif_lrn}. Each 3×3 matrix represents the suffix spelling system in use at a particular generation with shape represented along the rows and color represented along the columns, following the same 3×3 layout used in Fig.~\ref{stimuli}. The color-coding of these matrices indicates similarity in suffix form: Similar colors are used to represent similar suffixes, making it easier to see how the suffixes pattern with meaning.\footnote{The color-coding was generated independently for each chain by selecting $n$ evenly-spaced hues, where $n$ is the number of unique forms that emerged across the chain, and mapping these hues onto the suffixes in alphabetical order.} For example, the system at Generation~9 in Chain~D uses three spellings (<koe>, <ko>, and <co>) to express the shape dimension, yielding a horizontal stripes pattern in the matrix representation. We describe such a system as ``redundant'' because shape was consistently and reliably expressed by the stem, so the suffix spelling system that emerged in this case conveys no additional information---the suffix simple repeats whichever shape was marked by the stem. Redundant suffix systems are characteristic of the Transmission-only condition, with similar outcomes occurring in Chains~C, E, F, and~H. We also saw the emergence of fully transparent suffix spelling systems in Chains~A, B, I, and~J. Chain~I, for example, ultimately uses a single spelling, <coe>, to represent the /kəʊ/ sound; that is, the written suffix forms make no distinction between shapes or colors, just like the spoken language. Chain~G did not settle on a clear pattern, using <co>, <coe>, and <coh> somewhat interchangeably in the final generation, although there were some signs of a color-expressive system emerging in, for example, Generation~6, where yellow is consistently spelled <coh>, blue is consistently spelled <coe>, and pink uses both <co> and <coe>. The only other signs of color-expressive systems are H1, which was rapidly converted into a redundant system in subsequent generations, and J2, which ultimately degenerated towards transparency.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/dif_com.eps}}
	\vspace*{2pt}
	\caption{Results from the Transmission~+ Communication condition in Experiment~1 (differentiation). Each matrix shows the suffix spelling system in use at a particular generation (shape on the rows, color on the columns, as in Fig.~\ref{stimuli}). Chains are labeled K--T and generations are labeled 0–9 (0 is the randomly generated seed system). Each chain uses an independent color palette, with each color representing a particular suffix spelling; similar colors indicate similar spellings. Spellings in bold-italic are the generalizations on unseen items. There are some isolated examples of differentiation through implicit generalization (K5, M1, P5, S2) and explicit innovation (K9, L7, M7, O4, Q7).}
	\label{dif_com}
	\end{figure}

The results for the Transmission~+ Communication condition (Chains~K--T) are shown in Fig.~\ref{dif_com}. Like Transmission-only, degeneration to a single spelling by the ninth generation is a relatively common outcome (e.g., Chains~N, R, and~S and to a lesser extent Chains~L, M, and~O). In contrast, redundant, shape-expressive systems (as indicated by horizontal stripes) are relatively rare (e.g., K8, M6 and Q1--6). Instead, the presence of the communicative task appears to favor color-expressive systems, although these are far from common and often unstable. In particular, there are two main kinds of color-expressive system that emerge. The first are the generalization-based expressive systems: K5, M1, P5, and S2. In these cases, the participant tended to generalize their input in a way that is consistent with expressing color over shape. For example, in K5, pink was consistently spelled <ko> and blue was consistently spelled <co>. In the case of M1, color is expressed by the final vowel letters (<oe> for pink, <o> for yellow, and <oh> for blue) with the spelling of the /k/ sound conditioned on the stem (<buvic->, <zetik->, and <wopiq->). We note, however, that in all these cases the expressive system was not reciprocated by the participant's partner---resulting in low communicative accuracy---and not sustained or elaborated~on in subsequent generations.

The second type of color-expressive system is one that simply appropriates the expressive power of English to differentiate color: K9, M7, Q7, and to a lesser extent L7 and O4. In M7, for example, the participant added <r>, <g>, and <b> (presumably red, gold, blue) to the ends of the words, although the participant's partner only reciprocated the <g> spelling and seemingly failed to understand what was meant by <r> and <b>. In the case of Q7, the pair of participants added <r>, <o>, and <b> (presumably red, orange, blue) to communicate with high accuracy, but, although this system was retained into Generation~8, it started to disintegrate by Generation~9. Overall, in the five cases where English color letters were used, the systems did not really catch~on, perhaps because they retained spelling redundancy from the previous generation, resulting in unnecessarily complex suffix spellings that were too difficult to learn for subsequent generations. Q7, for example, uses <-coe->, <-co->, <-qo-> for shape plus <-r>, <-o>, <-b> for color. In addition to this handful of cases, there were a further four pairs of participants who used full English words as the suffix (typically <-pink> or <-red>, <-oran> or <-yell>, and <-blue>), but these pairs were excluded and replaced before iteration to the next generation.\footnote{In our preregistration, we stated that we would reject participants who used English color words. We allowed single letters because these were typically less transparent, especially in cases where the choice of letter (e.g., <g> or <r>) did not match how another person might have perceived the color (yellow vs.\@ gold vs.\@ orange or pink vs.\@ red). Indeed, as we see in the results, subsequent generations often did not pick~up~on the color letters (L9, M8, and Q9).} We mention these cases, however, because they are still examples of a conscious effort to differentiate---something that never happened in the Transmission-only condition.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/typology.eps}}
	\vspace*{2pt}
	\caption{Four primary systems of interest (or typological categories) arranged along the cost/complexity continuum. A holistic system uses a unique suffix spelling for each shape--color combination. An expressive system only expresses color. A redundant system only expresses shape (which is already conveyed by the stem). A degenerate system expresses nothing.}
	\label{typology}
	\end{figure}

To analyze more formally what types of system tend to emerge under different conditions, we first designated four suffix systems that are of particular interest: holistic, expressive, redundant, and degenerate. These four typological categories may be positioned along the simplicity--informativeness continuum in terms of the number of suffix forms they make use of and how these forms are conditioned on meaning, as illustrated in Fig.~\ref{typology}. The holistic\footnote{Our use of the term \textit{holistic} is slightly unusual here, owing to the fact that we are focused on the suffix level and not the word level. By holistic, we only mean that each shape--color combination has a unique suffix form, so the nine suffixes must all be learned as separate cases. We do not distinguish between truly holistic suffixes (nine unique suffixes with no structure in how they relate to each other) and compositional suffixes (nine unique suffixes that can be generated from compositional rules, as was the case in, for example, Q8).} system has nine unique forms each of which expresses a particular color--shape combination. This is complex to learn, but the suffix alone can pick out exactly one meaning. The expressive and redundant systems have three unique forms that express either color (expressive) or shape (redundant). These systems are easier to learn, but only the expressive system is fully informative (when acting in combination with the stem). The degenerate system uses just one suffix form. This makes it trivial to learn, but entirely uninformative. To classify a participant's output into one of these four typological categories, we computed which of the reference systems was most similar in terms of its information content. To do this, we formalized the systems as set partitions (i.e., a partitioning of the universe of meanings into disjoint subsets) with variation of information \parencite{Meila:2007} defined as the distance metric between any two such partitions.\footnote{Variation of information is a proper metric on set partitions, measuring the amount of information (in bits) that is lost and gained in the transformation of one partition into another. Under this metric, the holistic and degenerate systems are considered very dissimilar because they carry very different levels of information content. The expressive and redundant systems are also considered quite dissimilar because, although they carry the same amount of information, the information they carry is orthogonal (shape in the case of redundant, and color in the case of expressive).} A given participant's output system is then classified into one of the four typological categories based on whichever reference system is closest.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/typ_dist_dif.eps}}
	\vspace*{2pt}
	\caption{Upper panels show the typological distribution at each generation and in each condition across four typological categories: holistic (H; purple), expressive (E; blue), redundant (R, green), and degenerate (D; yellow). Bottom panels show posterior differences (mean and 90\% HDI) in the proportion of chains falling into each typological category; positive values signify greater use of a given category in the communicative condition.}
	\label{typ_dist_dif}
	\end{figure}

The typological distributions are plotted in Fig.~\ref{typ_dist_dif}, reveling the proportion of the 10 chains that fall into each typological category at each generation. In Transmission-only, the holistic systems used to initialize the chains are rapidly replaced by redundant systems by Generation~2. The dominance of the redundant category is then gradually eroded as the chains transition to degeneracy. In Transmission~+ Communication, there is initially a fairly even mix of all four kinds of system, but by Generation~9, degenerate systems tend to be most common. There is also a notable increase in holistic systems emerging in later generations; these are the cases of compositional suffixes that arose through the addition of English color letters on top of redundant suffix spellings (i.e., K9, M1, M7, and Q7). To quantify our uncertainty about these distributions and compare conditions statistically, we adopt a Dirichlet-Multinomial model of typological distribution $t$ for generation $i$,
	\begin{equation}
	\begin{aligned}
	   t_i & \sim \mathrm{Multinomial}(10, \theta_i) \\
	   \theta_i & \sim \mathrm{Dirichlet}([1, 1, 1, 1]),
	\end{aligned}
	\end{equation}
and compute the posterior difference in $\theta_i$ (the estimated proportion of chains falling into each category) across conditions: $\Delta(\theta_i) = \theta_{i_\mathrm{comm}} - \theta_{i_\mathrm{trans}}$. The means and 90\% HDIs of these posterior differences are plotted in the bottom panels of Fig.~\ref{typ_dist_dif}. In general, we find no clear difference in the number of expressive systems between the two conditions; however, we do find that redundant systems are less common in Transmission~+ Communication (in most generations), while holistic systems tend to be a little more common (most notably at Generation~7). Although there is not much evidence to suggest that Transmission~+ Communication strongly favors more informative systems (i.e., the expressive and holistic categories), it is interesting to note that the communicative condition did seem to disfavor the inexpressive redundant systems.

As a more general measure of how informative the written words are over time (i.e., when the stem and suffix act in combination), we adopt communicative cost \parencite{KempRegier:2012, Regier:2015, Kemp:2018}. This measure expresses how much information is lost, on average, when a speaker/writer attempts to convey a meaning to a listener/reader using some shared signaling system. If the system contains no ambiguities (all meanings are expressed by unique signals), communicative cost will be zero---that is, zero information will be lost during each attempt to communicate using that system. Communicative cost will take some larger value if the system contains ambiguity. It is given by
	\begin{equation}
	\sum_{m \in U} \mathrm{Pr}(m) ( -\log \mathrm{Pr}(m|s_m) ),
	\end{equation}
where $U$ is the universe of meanings that may be expressed, $\mathrm{Pr}(m)$ is the probability that a particular meaning would need to be expressed, and $\mathrm{Pr}(m|s_m)$ is the probability that a reader would infer meaning $m$ given that a writer produced signal $s$ for meaning $m$. In our case, $U$ is the set of nine alien objects, $\mathrm{Pr}(m)$ is set to $1/|U|$ (all objects need to be talked about with equal probability), and $\mathrm{Pr}(m|s_m)$ is given by $1/|M_s|$, where $M_s$ is the set of meanings labeled $s$ according to the system.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/cost_dif.eps}}
	\vspace*{2pt}
	\caption{\textbf{A} Communicative cost by generation under Transmission-only (dark green) and Transmission~+ Communication (orange). The thin lines show each individual transmission chain (with a small amount of added jitter). The thick lines and shaded regions show the mean and 90\% HDI of a polynomial regression model fit to the data. Generation~0 is not included in the model, since it does not represent participant data. The black dotted lines show the communicative cost of the spoken language for comparison. In general, the orthographic systems tend to become more costly over time, but less so under communicative pressure. \textbf{B} Posterior parameter estimates from the regression model, along with the posterior differences across conditions.}
	\label{cost_dif}
	\end{figure}

Fig.~\ref{cost_dif}A plots communicative cost over the nine generations of cultural transmission in each condition. The initial, generation-0 seed systems are all fully informative (zero cost) because each meaning is signaled by a unique arbitrary spelling. In both conditions, cost increases with generation, implying that the languages become less informative over time. This is to be expected in the Transmission-only condition, since there is only pressure from learnability, so the systems evolve to become simple and transparent. In contrast, this is not the expected result in Transmission~+ Communication, where the communicative pressure should promote the emergence of informative systems. This result does, however, align with our qualitative interpretation of the data, which suggested that participants in the communicative condition rarely produced differentiated spellings, be it through explicit innovation or implicit generalization. Interestingly, however, cost plateaus at a slightly lower level in Transmission~+ Communication, suggesting that there is some tendency for the systems to remain slightly more informative in the communicative condition.

To evaluate this argument statistically, we fit the following regression model:
	\begin{equation}
	\begin{aligned}
	   c_i & \sim \mathrm{Normal}(\mu_i, \sigma) \\
	 \mu_i & =    \alpha + \beta_1 g_i + \beta_2 g_i^2 + \beta_3 g_i^3 \\
	\sigma & \sim \mathrm{Exponential}(1),
	\end{aligned}
	\end{equation}
where $g$ is a centered generation number (i.e., Generations~1 through 9 are represented as~$-4$ through $+4$). The model states that the cost~$c$ of participant~$i$'s output system is drawn from a normal distribution, the mean of which is determined by a third-degree polynomial model. The intercept $\alpha$ corresponds to Generation~5 (the central generation), and the $\beta$ parameters determine the shape of the polynomial. $\sigma$ represents an across-participant noise term and is given a regularizing exponential prior. The priors on $\alpha$ and $\beta$ are not shown here, for brevity, but these parameters are estimated separately for each chain and pooled into global estimates with the use of Normal hyperpriors. For the complete model specification, see the Supplementary Material.

The posterior parameter estimates of the model are shown in Fig.~\ref{cost_dif}B. Essentially, the $\beta$ parameters do not differ substantially between conditions, but the $\alpha$ parameters do differ. This may be interpreted as the polynomial regression models adopting approximately the same shape but plateauing at different levels of cost. Additionally, we find that the $\alpha$ parameters themselves are lower than the cost of the spoken language (indicated by the dotted line), although it is unclear how this situation might unfold beyond Generation~9, especially in Transmission-only where the model predicts that cost will continue to increase further until all orthographic systems become aligned with the spoken language.

Finally, cost may be converted into the expected probability of communicative success\footnote{Note that we can compute this quantity regardless of which condition a participant was in. By expected probability of communicative success, we mean the level of success a pair of interlocutors could hypothetically achieve using a given system, regardless of whether that system emerged in the Transmission-only or Transmission~+ Communication condition. This also assumes that the hypothetical interlocutors have a perfect understanding of the same shared system.} with the equation $p=2^{-c}$. If we take Generation~5 as representative of the level at which cost plateaus, we find that estimated communicative success is $2^{-\alpha} = 40.8$\% (90\% HDI: [35.1, 46.3]) in Transmission-only and 50.6\% (90\% HDI: [45.9, 54.9]) in Transmission~+ Communication, representing an advantage of around 10 percentage points for the communicative condition. Overall, then, while the presence of the communicative pressure does appear to make a meaningful difference to the informativeness of the emergent systems, the levels of informativeness that were achieved were not especially impressive considering that a degenerate system would result in a communicative success rate of~33.3\% (i.e., chance-level performance), while an expressive system would result in a~100\% success rate, while also being fairly easy to learn.

\subsection{Summary}

Our first experiment tested whether informative, heterographic orthography could emerge through differentiation and whether it would emerge preferentially under communicative pressure. Overall, there was little evidence of systematic orthographic differentiation. Other than resorting to English, such differentiation could have been achieved in a number of ways, most obviously through the conditioning of spelling variation on meaning (e.g., <k> for pink, <c> for yellow, and <q> for blue), but also through less obvious strategies such as the use of length (e.g., <o> for pink, <oo> for yellow, and <ooo> for blue) or the use of arbitrary silent letters (e.g., <a> for pink, <b> for yellow, and <c> for blue). Such differentiation was, however, not very forthcoming, even under communicative pressure. Instead, when spelling variation was conditioned on anything, it was usually conditioned on shape, resulting in redundant suffix spellings. This result is in stark contrast to most prior experimental iterated learning studies, in which informative, compositional systems do typically emerge, especially under communicative pressure \parencite[e.g.,][]{Kirby:2015}.

So, what was different about the present experiment? The primary difference was the presence of a spoken language that is decidedly \textit{not} informative about one of the dimensions. Indeed, the point of our experiment is to see whether orthography can resist phonology under sufficient pressure for informativeness. The presence of homophonous suffix forms acts as a cue to participants that the language itself does not mark color and that, therefore, the orthography should also not mark color. Faced with variation that could be conditioned on either dimension, learners appear to rule out the possibility that it might be conditioned on color, since such a hypothesis would be in conflict with the spoken language. As a result, any variation in spelling comes to be associated with particular stems, resulting in the emergence of redundant suffix spellings that serve no real purpose. A rough analog of this redundancy can be found in English in the spelling of /-ʃən/ (<cian>, <sion>, <ssion>, or <tion>), which is largely conditioned on the stem (e.g., \textit{magician}, \textit{expulsion}, \textit{transmission}, and \textit{station}).

It is particularly surprising that participants chose not to innovate novel systems in the communicative condition, despite their being a clear financial incentive to do so. Why might this be the case? Firstly, although innovation was not expressly forbidden in the instructions, it was perhaps suggested by the overall structure of the experiment that participants were expected to use the orthography learned during training. Participants who adopted this interpretation may have feared diverging from what they had learned, either because they themselves wanted to avoid breaking the rules or because they did not know whether their partner would feel comfortable breaking the rules. Secondly, even if participants were willing to diverge, they were presented with an alignment problem: How should they decide what system of color marking to use when there is no channel through which they can agree a system with their partners? They can only use trial-and-error, and even then, the process is complicated by participants simultaneously adapting to each other's systems. These are dynamics that also exist in the real world, especially in writing where the reader and writer are separated in time and space. Deviation from spelling norms is often considered socially unacceptable or uneducated, even when novel spellings are advantageous (e.g., the non-standard spelling <thru> has struggled for decades to make any headway against <through>). Even under more progressive attitudes to change and innovation, the problem of achieving community buy-in remains, especially when the change is not imposed by top-down authority: If someone decided to start using the spelling <banque> to differentiate the financial institution from river banks, would anyone take that spelling seriously or even understand the distinction? Differentiation is hard in the written modality.

%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%

\section{Experiment 2}

We now turn our attention to the conservation model of heterographic homophones: Given that an informative system already exists (both in speech and in writing), does that informative system persist in writing even after the spoken language has degenerated into homophony? And, importantly, does this happen preferentially in the presence of communicative pressure?

\subsection{Methods}

The methods were identical to Experiment~1 with two exceptions: The artificial language used to seed each chain started out fully compositional (in both its spoken and written forms), and two sound mergers were artificially induced during cultural transmission, resulting in the spoken forms of the suffixes becoming increasingly homophonous and uninformative over time.

\subsubsection{Participants}

The experiment was completed by 297 native-English participants recruited through Prolific. The payment and bonusing scheme was identical to Experiment~1 (median bonus: £0.78). The median completion time was 15~m with a median hourly rate of £7.99 (£10.68 including bonus). 17 participants were excluded because they (or their partners) failed the auditory attention checks (15) or used English color words (2). A further 10 participants were lost to communication-game pairing failures. Like Experiment~1, the final dataset comprises 270 participants: 90 in the Transmission-only condition (10 chains of 9 participants) and 180 in the Transmission~+ Communication condition (10 chains of 9 pairs of participants).

\subsubsection{Stimuli}

The alien objects and word stems were identical to Experiment~1. Unlike Experiment~1, however, the transmission chains were seeded with a fully compositional language that used three distinct suffixes to systematically express each of the colors. A separate set of suffixes was created for each chain by concatenating a randomly drawn consonant from \{/f/, /s/, /ʃ/\} and a randomly drawn vowel from \{/ə/, /ɛɪ/, /əʊ/\}, both without replacement. For example, one chain might use the suffixes /fəʊ/, /ʃə/, /sɛɪ/ to represent the colors pink, yellow, and blue, while another chain might use /sə/, /fɛɪ/, /ʃəʊ/ for those colors. The initial orthographic system was transparent and based on the following phoneme–grapheme mapping: \{/f/→<f>, /s/→<s>, /ʃ/→<x>, /ə/→<a>, /ɛɪ/→<ei>, /əʊ/→<oe>\}. The suffixes were designed to be distinctive (and therefore easy to memorize and associate with colors), but also similar enough to (mostly) allow for somewhat plausible sound mergers and result in somewhat plausible spellings following sound merger (e.g., it is plausible that /f/ might supplant /s/ or /ʃ/ in speech or that /ʃ/ might be spelled <s> or <x> in writing). We attempted to achieve this balance by combining consonants that are very similar with vowels that are very dissimilar, while also avoiding reuse of any sounds present in the stems. The spoken forms were synthesized using the Apple text-to-speech synthesizer (Moira voice).

\subsubsection{Sound change}

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/sound_change.eps}}
	\vspace*{2pt}
	\caption{Examples of the spoken suffixes in Experiment~2. During Epoch~I, color is represented by three distinct spoken suffixes. During Epoch~II, two of the suffixes are homophonous, reducing the informativeness of the spoken language. During Epoch~III, the spoken form of the language makes no color distinction.}
	\label{sound_change}
	\end{figure}

Each transmission chain was run for nine generations, which were divided into three epochs. During Epoch~I (Generations~1 to~3), the three spoken suffixes were distinct (as described above), allowing the spoken language to express all three colors without ambiguity. During Epoch~II (Generations~4 to~6), the spoken language had two distinct suffix forms, reducing its informativeness. During Epoch~III (Generations~7 to~9), all three spoken suffixes were homophonous, just as in Experiment~1, making the spoken language entirely uninformative about color. This was achieved through two sets of sound changes, the first occurring in the transition from Epoch~I to~II and the second occurring in the transition from Epoch~II to~III. An example is illustrated in Fig.~\ref{sound_change}. In the first sound change, two of the spoken suffix forms were chosen at random and the consonant from one (chosen at random) was paired with the vowel from the other, resulting in a new suffix form that replaced the original two. In the second sound change, the remaining two suffix forms were merged in the same way, resulting in full homophony. Crucially, the spellings did not automatically change following a sound change event; rather, the orthographic system was free to adapt (or not) in response to the sound changes. Note also that an individual participant did not directly experience the sound changes; a Generation~4 participant, for example, would always hear Epoch~II sounds, while observing spellings produced by a Generation~3 participant (presumably representing the Epoch~I sounds). In reality, of course, sound change is more gradual, with individual speakers experiencing both outgoing and incoming spoken forms within their lifetimes.

\subsection{Results}

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/con_lrn.eps}}
	\vspace*{2pt}
	\caption{Results from the Transmission-only condition in Experiment~2 (conservation). Each matrix shows the suffix spelling system in use at a particular generation (shape on the rows, color on the columns, as in Fig.~\ref{stimuli}). Chains are labeled A--J and generations are labeled 0–9 (0 is the randomly generated seed system). Each chain uses an independent color palette, with each color representing a particular suffix spelling; similar colors indicate similar spellings. Spellings in bold-italic are the generalizations on unseen items. The final systems are characterized by the conservation of form without the conservation of expressivity.}
	\label{con_lrn}
	\end{figure}

The results for all ten chains (labeled A--J) in the Transmission-only condition are shown in Fig.~\ref{con_lrn}. It is immediately clear that color-expressive suffixes (as indicated by vertical stripes) are maintained fairly reliably through the first epoch; perfectly in the case of Chains~A, B, D, F, and~I, and with some errors in the other five chains. Some of these errors are very minor, such as the use of <sha> instead of <xa> for one item in H3, while others are more catastrophic, such as the early loss of the <sei>/<xa> distinction in Chain~J. Generation~4 represents the first real test of the orthographic systems in the face of sound change, and, in most cases, the Generation~3 systems are preserved quite faithfully (Chains~A, B, D, E, H, and~I), but by the end of the epoch (Generation~6), many have degenerated into redundant systems that encode shape (Chains~B, G, and~J) or transparent systems that mirror the Epoch~II pattern of homophony (Chains~C, E, F, and~I). These processes continue into Epoch~III and by the ninth generation, all systems have become degenerate, redundant, or some mixture of the two. The one exception is Chain~D, whose original spellings---<sei>, <fa>, <xoe>---were conserved perfectly through to the final generation with only one generalization error in Generation~8, which was quickly reverted in Generation~9.

Like Experiment~1, redundant systems are characteristic of the Transmission-only condition, especially in Epoch~III. As the spoken suffixes become more homophonous, the variant spellings are increasingly conditioned on shape rather than color, perhaps because the spoken language signals to learners that the language does not mark color, so they rationalize the system as three words with idiosyncratic spellings. Interestingly, however, all chains exhibited conservation of spelling \textit{form}, even if the way in which spelling was conditioned on meaning was lost. For example, Chain~A ultimately represents the sound /fə/ with the spellings <foe> and <xa>, spellings that are internally inconsistent and contrary to standard uses of the Roman alphabet, but which trace their origins back to the original seed orthography. Overall, then, the Transmission-only condition in Experiment~2 is characterized by the conservation of spelling \textit{form} without conserving how form patterns with meaning.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/con_com.eps}}
	\vspace*{2pt}
	\caption{Results from the Transmission~+ Communication condition in Experiment~2 (conservation). Each matrix shows the suffix spelling system in use at a particular generation (shape on the rows, color on the columns, as in Fig.~\ref{stimuli}). Chains are labeled K--T and generations are labeled 0–9 (0 is the randomly generated seed system). Each chain uses an independent color palette, with each color representing a particular suffix spelling; similar colors indicate similar spellings. Spellings in bold-italic are the generalizations on unseen items. Five chains (O, P, Q, R, T) remain fully expressive into the final epoch, in most cases conserving the original forms.}
	\label{con_com}
	\end{figure}

The results for the Transmission~+ Communication condition (Chains~K--T) are shown in Fig.~\ref{con_com}. Like Transmission-only, the seed systems are mostly maintained faithfully through Epoch~I; perfectly in the case of Chains~K, L, N, O, and~T, and with some errors in the other five chains, although some of these errors are non-destructive changes, such as <x> being replaced with <sh> in R1. Several systems were then maintained through Epoch~II, notably Chains~O, P, R, and~T, and, in one case, through to the end of Epoch~III (Chain~R, albeit with the original <xei> spelling replaced with <shei>). The Chain~O system was preserved up to Generation~7, Chain~P was maintained up to Generation~7 and almost to Generation~9 with two modifications (<sha> instead of <xa> and <fa> instead of <fei>), and Chain~T was conserved faithfully up to Generation~8. The final form of Chain~Q was also fully expressive, albeit through a combination of both conservation and differentiation: The <fei> form was conserved from the seed orthography, the <oxie> spelling appears to derive from a misremembering of <xoe> (partial conservation), and the <fe> spelling, which began as a typo introduced in Q6, seems to have been generalized across the blue items in Q7, perhaps to differentiate them from the yellow items (indeed, the participant's partner made the same generalization). A similar case of differentiation might also have occurred in L9, where the <sol> spelling (originally a typo on <so>) was generalized across the yellow items, resulting in a semi-expressive system (although the participant's partner generalized the <sol> spelling across shape).

Unlike Experiment~1, no participant pairs attempted to use English color letters and there was only one case of a pair using English color words (Generation~9 of Chain~K), although this pair was excluded and replaced (this generation was the only case in which the training input was fully degenerate, which would have resulted in a strong pressure to find a communicative solution in the form of English). Presumably, the general conservation of expressive spelling in Experiment~2 negated the need to innovate novel systems.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/typ_dist_con.eps}}
	\vspace*{2pt}
	\caption{Typological distribution at each generation and in each condition across four typological categories: holistic (H; purple), expressive (E; blue), redundant (R, green), and degenerate (D; yellow). At generation~0, all 10 chains in both conditions use an expressive system (by design). In Transmission-only, the systems tend to transition to the redundant or degenerate patterning, while in Transmission~+ Communication, they tend to remain expressive for longer and disfavor the redundant pattern. Posterior difference (mean and 90\% HDI) in the proportion of chains falling into each typological category. Positive values signify greater use of a given typological category in the communicative condition (and vice versa for negative values).}
	\label{typ_dist_con}
	\end{figure}

Fig.~\ref{typ_dist_con} plots the typological distributions by generation and condition. Initially, all systems are expressive, but the dominance of this typological category is gradually eroded over time, particularly during the second and third epochs once the spoken forms had become homophonous. Notably, however, the loss of expressive systems appears to be slower in Transmission~+ Communication, and redundant systems were also less popular under communicative pressure. The clearest difference between conditions comes at Generation~7, at which point there were more expressive systems and fewer redundant systems in the communicative condition.

	\begin{figure}
	\makebox[\textwidth][c]{\includegraphics[]{figs/cost_con.eps}}
	\vspace*{2pt}
	\caption{\textbf{A} Communicative cost by generation under Transmission-only (dark green) and Transmission~+ Communication (orange). The thin lines show each individual transmission chain (with a small amount of added jitter). The thick lines and shaded regions show the mean and 90\% HDI of a regression model fit to the data. Generation~0 is not included in the model, since it does not represent participant data. The black dotted lines show the communicative cost of the spoken language, which increases with epoch. In general, the orthographic systems tend to remain more informative under communicative pressure. \textbf{B} Posterior parameter estimates from the regression model, along with the posterior differences across conditions.}
	\label{cost_con}
	\end{figure}

Fig.~\ref{cost_con}A plots communicative cost. During Epoch~I, there is a shallow increase in cost in both conditions, representing a baseline level of noise when the spoken language is fully informative. During Epoch~II, there is a steep rise in cost in the Transmission-only condition, but a comparatively shallow rise in the Transmission~+ Communication condition, presumably due to the communicative pressure, which slows the rate at which the systems become transparent or redundant. In Epoch~III, cost plateaus in the Transmission-only condition because the written languages have become sufficiently simple to be reliably learned and reproduced with little change. In Transmission~+ Communication, by contrast, cost continues to increase at much the same rate as in Epoch~II.

To evaluate statistically the effect of generation on cost across the two conditions, while also taking epoch into account, we fit the following regression model:
	\begin{equation}
	\begin{aligned}
	   c_i & \sim \mathrm{Normal}(\mu_i, \sigma) \\
	 \mu_i & =    \alpha_{\mathrm{EPOCH}[i]} + \beta_{\mathrm{EPOCH}[i]} g_i \\
	\sigma & \sim \mathrm{Exponential}(1),
	\end{aligned}
	\end{equation}
where $g$ is an epoch-centered generation number (i.e., for Generations~1, 4, and~7, $g=-1$, for Generations~2, 5, and~8, $g=0$, and for Generations~3, 6, and~9, $g=1$). Overall, the model states that the cost~$c$ of participant~$i$'s output system is drawn from a normal distribution, the mean of which is determined by an epoch-specific linear model (i.e., each epoch has an independent intercept $\alpha$ and slope $\beta$). $\sigma$ represents an across-participant noise term and is given a regularizing exponential prior. The priors on $\alpha$ and $\beta$ are not shown here, for brevity, but these parameters are estimated separately for each chain and pooled into global estimates with the use of Normal hyperpriors. For the complete model specification, see the Supplementary Material.

The posterior parameter estimates are shown in Fig.~\ref{cost_con}B. The $\alpha$ parameters, which represent the middle generation of each epoch (Generations~2, 5, and 8 respectively), reveal the gradual divergence of the conditions: By Epoch~III, the difference is $-0.43$~bits (90\% HDI: [$-0.83$, $-0.04$]), meaning that the communicative condition is about half a bit more informative. Further, after Epoch~I, the orthographic systems tend to remain more informative than the spoken language, especially under the communicative condition. The $\beta$ parameters, which represent the slopes within each of the three epochs are not conclusively different between conditions, although there is some evidence to suggest that the slope is shallower in the communicative condition during Epoch~II (e.g., 89\% of the posterior probability mass on $\Delta(\beta_2)$ is below zero). In the case of $\beta_3$ the effect is reversed, with Transmission~+ Communication having a greater slope. Note, however, that this appears to be because cost has, by this stage, plateaued in the Transmission-only condition, so the increased homophony pressure introduced in Epoch~III has little effect in Transmission-only, but continues to exert an effect in Transmission~+ Communication.

At Generation~7---the point at which the two conditions differ most substantially---the difference in cost is $-0.6$~bits (90\% HDI: [$-1.04$, $-0.14$]). Converting cost to the probability of communicative success, we obtain an estimate of 46.5\% (90\% HDI: [38.1, 54.7]) in Transmission-only and 70.9\% (90\% HDI: [52.8, 87.8]) in Transmission~+ Communication, representing an advantage of around 25 percentage points for the communicative condition. The equivalent value for the spoken language at Generation~7 is 33.3\% (i.e., because in Epoch~III all three colors are expressed homophonously). Thus, not only are the two conditions different from each other, but also the two conditions are more informative than the spoken language, especially under communicative pressure.

\subsection{Summary}

Experiment~1 asked whether an informative, heterographic orthography may be \textit{created} de~novo under homophony. Experiment~2, by contrast, asked whether an informative, heterographic orthography can simply be \textit{maintained}, even under the same levels of homophony encountered in Experiment~1. In the Transmission-only condition, only one chain (Chain~D) remained expressive into the fully homophonous Epoch~III, while in Transmission~+ Communication, five chains (O,~P, Q, R, and~T) remained expressive, albeit not necessarily all the way to Generation~9. The fact that expressive spellings persisted longer and across more chains under communicative pressure suggests that an informative orthography---despite running contrary to the spoken language---may be maintained when it serves a useful purpose. That being said, the fact that informativeness could, in~principle, be maintained without communicative pressure (most notably in Chain~D) suggests that a strong communicative pressure is not a strictly necessary condition for conservation: Learning alone can, to a limited extent, maintain informative heterography.

Many of the chains did, however, eschew informativeness entirely in favor of greater transparency, and the inevitable long-term consequence for all chains appears to be degeneracy. This is to be expected under Transmission-only, where the systems are adapting under learnability pressure, but is somewhat surprising in Transmission~+ Communication. Our findings ultimately suggest that, in the long term, (alphabetic) orthographic systems favor the faithful encoding of speech over the useful encoding of meaning, although there may exist short windows of time during which informative heterography can resist the spoken language. This points to a preference that readers might possess---those accustomed to an alphabetic writing system, at least---in which the primary goal of writing is seen as representing speech, not representing meaning directly. Interestingly, although participants were quite resistant to encoding into writing something that is not encoded in speech, they were---at the same time---content to conserve spelling forms that were internally inconsistent and unusual. Tradition has a powerful hold over writing systems.

%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%

\section{Discussion}

Writing---particularly writing in alphabetic scripts---is usually viewed as a visual representation of speech: Each visual letter corresponds to a sound. However, writing systems are known to diverge from their spoken counterparts in various ways. The insertion of spacing between words, for example, is almost ubiquitous across alphabetic writing systems, even though no such spacing exists between words in speech. It seems likely that graphic innovations such as these exist because they confer some benefit that is not required in the spoken modality. In the case of spacing, for example, the separation of words into discrete chunks presumably aids in the targeting and extraction of visuo-linguistic information---constraints that do not exist in the auditory modality. In principle, the same may be true of spelling: Words may be spelled in ways that diverge from the spoken language, if that divergence confers some benefit in reading.

We tested this hypothesis by focusing on the particular case of heterographic homophones---words that sound the same but which are spelled differently. Heterographic homophones permit the written language to be more informative than the spoken language. The written forms <knight> and <night>, for example, convey a distinction in meaning that cannot be conveyed in speech without supplying additional information, either in the form of context or paraphrasis. In English, at~least, which has a tendency to preserve etymology in spelling, the heterographic homophones have typically been viewed as accidents of history: \textit{right}, \textit{rite}, and \textit{write} are spelled differently because their spellings have been inherited from the conventions of the languages of origin (Old English \textit{riht}, Latin \textit{ritus}, and Old English \textit{writan}). Nevertheless, examples of explicit efforts to differentiate the spelling of homophones do exist, such as x, y, z, and, moreover, the etymological cases of heterography may, in part, owe their own success to the differentiation they inadvertently provide.

We put these ideas to the test by experimentally simulating the cultural evolution of orthography. In Experiment~1, we focused on the differentiation mechanism: Might variant spellings be used to differentiate meanings that are otherwise identical in speech? If so, we would expect levels of spelling differentiation to be greater when there is greater pressure for disambiguation, which we induced through the addition of a communication game. Importantly, the initial randomly generated orthographic systems that we used to seed the transmission chains contained high variation---that is, multiple ways of spelling the same sound. We included this variation because, for the differentiation mechanism to be viable, the writing system has to be receptive to spelling variation. A writing system that does not permit a one-to-many phoneme-to-grapheme mapping would not be capable of differentiating homophonous words. Only when variation is permitted and available can variant spellings be conditioned on meaning.

Although we found evidence to suggest that the written forms were slightly more informative than the corresponding spoken forms, systematic differentiation was, on the whole, rare, unstable, and fleeting, be it through implicit generalization of the supplied variants or explicit innovation of new variants. This is \textit{not} because participants were unable to learn variant spellings; indeed, in many cases variant spellings were stable but conditioned on shape rather than color. Nor was it because participants were unable to learn a system of color marking that is not expressed in the spoken language; we know from Experiment~2 that participants can learn such systems. Rather, it seems that, in Experiment~1, differentiation could not get of the ground because of two main factors: First, participants seem to have a strong bias against representing a meaning in writing when that meaning is not represented in speech. Put differently, participants seem to view spelling as subordinate to speech, so when asked to type in a word for a pink pentagon called /buvɪkəʊ/ to send to their partner, their natural inclination is to type a sequence of graphemes that transparently reflects that sequence of phonemes, not the meaning that those phonemes are attempting to express. Secondly, even when participants appreciate the need to differentiate the written forms to be successful in the task, they often appear reluctant or unable to do so, perhaps because they view the written language as immutable or because the problem of aligning with a partner---even in a~synchronous setting---is too difficult to overcome. It is notable, for example, that in the communication games, many attempts to differentiate with English color letters were not reciprocated.

In Experiment~2, we tested a different mechanism by which orthographies may end up possessing additional informativeness beyond that of the spoken form of the language: conservation. Under this mechanism, an expressive orthography does not emerge to fulfill a role but rather it is simply maintained from an earlier form of the spoken language that was expressive. This parallels what we know about many of the heterographic homophones in English, which arose as byproducts of either the preservation of etymology or phonological changes that were never assimilated into the written forms \parencite{Berg:2021}. That is not to say, necessarily, that this form of heterography is entirely an accident of history; rather, such spellings may, at least in part, have been preserved precisely because of the informativeness they provide in reading. If correct, this would predict that expressive orthography should be preserved preferentially under communicative pressure.

Our findings showed that informative heterography may be conserved more frequently and for longer periods under communicative pressure for disambiguation. There are two important caveats, however. First, we found that cultural transmission alone---that is, blind learning and reproduction---will result in at~least some conservation, not only in form but also in the conditioning of form on meaning. Cases such as Chain~D correspond to the ``accident of history'' explanation: Expressive orthography is preserved not because it serves any useful purpose (recall that in Transmission-only there is no functional need for the language to be informative), but because participants are simply reproducing what they learned, and what they learned has not (yet) placed a significant enough burden on learning for simplification to kick~in. The additional level of conservation that occurs in Transmission~+ Communication corresponds to the exaptation/repurposing explanation; that is, expressive orthography that originally served one purpose (representing speech) is maintained for a new purpose (representing meaning directly). The second caveat is that, in the long term, it appears that transparency will ultimately win the day, even under communicative pressure. Chain~O, for example, went from expressive to degenerate in two generations under full homophony pressure, and it seems likely that all chains will ultimately undergo the same transformation eventually. Informative heterography that arises through conservation is but a temporary oasis on the march toward transparency.

It is important to note, at this point, that the two experiments cannot be compared directly, although we made every effect to keep the two as close as possible. Fundamentally, participants---or more generally, the evolutionary systems---are being asked to do something quite different across the two experiments: create in Experiment~1 and maintain in Experiment~2. The demands of these two tasks are different, and one task or the other may be better suited to our experimental paradigm. Similarly, we cannot draw direct conclusions from the relative strength of the results across the two experiments. For example, although the difference between conditions was perhaps a little more dramatic in Experiment~1, this does not imply that the differentiation model is a better candidate mechanism; it suggests, rather, that communicative pressure is more essential for the differentiation mechanism to operate. Nevertheless, our experiments do serve to highlight the comparative difficulties involved in differentiation vs.\@ conservation. For differentiation to operate, numerous hurdles must be overcome: participants must recognize the problem they are faced with, they must be willing to introduce and tolerate spelling variation, they must able to align with an interlocutor separated in time and space, and they must be willing to rebel against their input. Conserving an expressive orthography is, by comparison, a relatively simple affair. If conservation is a simpler (and therefore, by Occam's razor, more plausible) mechanism in our experimental setup, then presumably the same is true of the real world.

Of course, all of these conclusions must be interpreted within the limitations of our studies; our experiments are, after all, highly simplified simulacra of the relevant real-world processes. Other than the general scaling-down of the x to an experimentally tractable

Communication game not a good model of written communication. link to texting which is written but near-synchronous.

color morphs a bit weird

do not consider top-down authority

unclear whether disambiguation matters in the real world. context usually makes things clear.

Nevertheless, even very small biases can be amplified by cultural transmission.

We foresee two interesting directions in which this work might be taken in the future. The first is a relatively simple extension of the present experiments. By removing the spoken forms from the test/communication~game phase of the experiment, we can simulate something closer to a society of silent readers; that is, a society wherein the spoken form is only salient during learning and less salient once the users switch to using the language primarily through the written modality. This minor change might have a dramatic effect on the results: Participants may be less inclined to simply transcribe how a word sounds and more inclined to differentiate and conserve. Such an experiment is not without pitfalls, however. Notably, our reason for providing the spoken form at the test stage was to be sure that participants were aware of the homophony; our experiment aims to test whether participants can overcome homophony, so it was vital to be sure that participants were indeed aware of it. Thus, this change to the experimental design might be open to concerns that, at the test stage, participants might simply forget everything about the pronunciation, such that the experiment becomes equivalent to prior iterated learning studies---that is, it becomes focused on the emergence of linguistic structure in general, not orthographic structure that runs contrary to the spoken language.

A second, more general direction in which this work could be taken would be to explore how well the differentiation and conservation models of heterography fit with historical data on language change. For example, by tracing the development of the heterographic homophones in English, we might be able to determine which of the models offer a better fit to the data, or indeed classify different cases as being derived through the differentiation or conservation mechanism. Perhaps the best way to proceed here would be to develop formal models of these two mechanisms and test them against diachronic data on the pronunciation and spelling of words that ultimately become heterographic.

As a final note, heterography is often derided as a source of unnecessary complexity that language learners must contend with. However, our results show that, when the communicative need for disambiguation is strong enough, language users may prefer to tradeoff some simplicity for greater informativeness in the writing system. Perhaps, then, efforts to reform and simplify spelling may, to some extent, be a little hasty, and more recognition should be given to the benefits that opaque orthographies provide in the context of reading.

%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%

\section{Acknowledgments}

\noindent This work was funded by a grant from the Leverhulme Trust.

\printbibliography

%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%~%

\clearpage

\section{Appendix}

\subsection{Experiment 1 regression model}

	\begin{equation}
	\begin{aligned}
	   c_i                   & \sim \mathrm{Normal}(\mu_i, \sigma) \\
	 \mu_i                   & = \alpha_{\mathrm{CHAIN}[i]} + \beta_{1,\mathrm{CHAIN}[i]} g_i + \beta_{2,\mathrm{CHAIN}[i]} g_i^2 + \beta_{3,\mathrm{CHAIN}[i]} g_i^3 \\
	\sigma                   & \sim \mathrm{Exponential}(1) \\
	\alpha_{\mathrm{CHAIN}}  & \sim \mathrm{Normal}(\alpha_\mu, \alpha_\sigma) \\
	\beta_{j,\mathrm{CHAIN}} & \sim \mathrm{Normal}(\beta_{j_\mu}, \beta_{j_\sigma}) \\
	\alpha_\mu               & \sim \mathrm{Normal}(1, 1) \\
	\alpha_\sigma            & \sim \mathrm{Exponential}(1) \\
	\beta_{j_\mu}            & \sim \mathrm{Normal}(0, 1) \\
	\beta_{j_\sigma}         & \sim \mathrm{Exponential}(1)
	\end{aligned}
	\end{equation}

\subsection{Experiment 2 regression model}

	\begin{equation}
	\begin{aligned}
	   c_i                                 & \sim \mathrm{Normal}(\mu_i, \sigma) \\
	 \mu_i                                 & = \alpha_{\mathrm{EPOCH}[i],\mathrm{CHAIN}[i]} + \beta_{\mathrm{EPOCH}[i],\mathrm{CHAIN}[i]} g_i \\
	\sigma                                 & \sim \mathrm{Exponential}(1) \\
	\alpha_{\mathrm{EPOCH},\mathrm{CHAIN}} & \sim \mathrm{Normal}(\alpha_{\mathrm{EPOCH}_\mu}, \alpha_{\mathrm{EPOCH}_\sigma}) \\
	\beta_{\mathrm{EPOCH},\mathrm{CHAIN}}  & \sim \mathrm{Normal}(\beta_{\mathrm{EPOCH}_\mu}, \beta_{\mathrm{EPOCH}_\sigma}) \\
	\alpha_{\mathrm{EPOCH}_\mu}            & \sim \mathrm{Normal}(0, 1) \\
	\alpha_{\mathrm{EPOCH}_\sigma}         & \sim \mathrm{Exponential}(1) \\
	\beta_{\mathrm{EPOCH}_\mu}             & \sim \mathrm{Normal}(0, 1) \\
	\beta_{\mathrm{EPOCH}_\sigma}          & \sim \mathrm{Exponential}(1)
	\end{aligned}
	\end{equation}

\clearpage

\begin{table}
\begin{center}
\begin{threeparttable}
\caption{Experiment 1 posterior parameter estimates}
\footnotesize
\begin{tabular}{lrrrrrr}
\toprule
 &   Mean &     SD &  HDI (5\%) &  HDI (95\%) &  ESS &  $\hat{R}$ \\
\midrule
\multicolumn{7}{c}{Transmission-only} \\
$\alpha_\mu$  &  1.298 &  0.123 &   1.098 &    1.498 &   55213 &    1.0 \\
$\beta_{1_\mu}$ &  0.003 &  0.030 &  -0.048 &    0.051 &   19478 &    1.0 \\
$\beta_{2_\mu}$ & -0.021 &  0.008 &  -0.034 &   -0.009 &   49158 &    1.0 \\
$\beta_{3_\mu}$ &  0.007 &  0.002 &   0.004 &    0.010 &   13927 &    1.0 \\
$\sigma$    &  0.192 &  0.020 &   0.160 &    0.225 &   26492 &    1.0 \\
\midrule
\multicolumn{7}{c}{Transmission + Communication} \\
$\alpha_\mu$  &  0.985 &  0.079 &   0.854 &    1.112 &   22089 &    1.0 \\
$\beta_{1_\mu}$ &  0.003 &  0.059 &  -0.091 &    0.100 &   19115 &    1.0 \\
$\beta_{2_\mu}$ & -0.017 &  0.007 &  -0.030 &   -0.005 &   15955 &    1.0 \\
$\beta_{3_\mu}$ &  0.004 &  0.004 &  -0.002 &    0.010 &   12773 &    1.0 \\
$\sigma$    &  0.370 &  0.034 &   0.313 &    0.425 &   27848 &    1.0 \\
\midrule
\multicolumn{7}{c}{Difference} \\
$\Delta(\alpha_\mu)$    & -0.313 &  0.147 &  -0.551 &   -0.072 &   38539 &    1.0 \\
$\Delta(\beta_{1_\mu})$   & -0.000 &  0.066 &  -0.109 &    0.107 &   19664 &    1.0 \\
$\Delta(\beta_{2_\mu})$   &  0.004 &  0.011 &  -0.014 &    0.022 &   24478 &    1.0 \\
$\Delta(\beta_{3_\mu})$   & -0.003 &  0.004 &  -0.010 &    0.003 &   12532 &    1.0 \\
$\Delta(\sigma)$    &  0.178 &  0.040 &   0.112 &    0.242 &   28446 &    1.0 \\
\bottomrule
\end{tabular}
\label{exp1_parameter_estimates}
\end{threeparttable}
\end{center}
\end{table}

\clearpage

\begin{table}
\begin{center}
\begin{threeparttable}
\caption{Experiment 2 posterior parameter estimates}
\footnotesize
\begin{tabular}{lrrrrrr}
\toprule
 &   Mean &     SD &  HDI (5\%) &  HDI (95\%) &  ESS &  $\hat{R}$ \\
\midrule
\multicolumn{7}{c}{Transmission-only} \\
$\alpha_{1_\mu}$ &  0.099 &  0.066 &  -0.009 &    0.205 &   50911 &    1.0 \\
$\alpha_{2_\mu}$ &  0.530 &  0.134 &   0.320 &    0.748 &   58969 &    1.0 \\
$\alpha_{3_\mu}$ &  1.113 &  0.155 &   0.869 &    1.370 &   55884 &    1.0 \\
$\beta_{1_\mu}$ &  0.048 &  0.044 &  -0.021 &    0.121 &   32489 &    1.0 \\
$\beta_{2_\mu}$ &  0.216 &  0.074 &   0.096 &    0.336 &   47941 &    1.0 \\
$\beta_{3_\mu}$ & -0.000 &  0.040 &  -0.065 &    0.065 &   32801 &    1.0 \\
$\sigma$      &  0.145 &  0.018 &   0.116 &    0.173 &   15490 &    1.0 \\
\midrule
\multicolumn{7}{c}{Transmission + Communication} \\
$\alpha_{1_\mu}$ &  0.052 &  0.042 &  -0.018 &    0.118 &   30904 &    1.0 \\
$\alpha_{2_\mu}$ &  0.297 &  0.102 &   0.138 &    0.469 &   56334 &    1.0 \\
$\alpha_{3_\mu}$ &  0.679 &  0.184 &   0.386 &    0.981 &   63206 &    1.0 \\
$\beta_{1_\mu}$ &  0.033 &  0.044 &  -0.040 &    0.102 &   18241 &    1.0 \\
$\beta_{2_\mu}$ &  0.093 &  0.073 &  -0.027 &    0.209 &   39754 &    1.0 \\
$\beta_{3_\mu}$ &  0.167 &  0.121 &  -0.033 &    0.359 &   53767 &    1.0 \\
$\sigma$      &  0.175 &  0.020 &   0.141 &    0.207 &   17902 &    1.0 \\
\midrule
\multicolumn{7}{c}{Difference} \\
$\Delta(\alpha_{1_\mu})$     & -0.047 &  0.079 &  -0.178 &    0.077 &   40911 &    1.0 \\
$\Delta(\alpha_{2_\mu})$     & -0.233 &  0.168 &  -0.504 &    0.040 &   55907 &    1.0 \\
$\Delta(\alpha_{3_\mu})$     & -0.434 &  0.241 &  -0.825 &   -0.037 &   57860 &    1.0 \\
$\Delta(\beta_{1_\mu})$     & -0.016 &  0.062 &  -0.118 &    0.084 &   23610 &    1.0 \\
$\Delta(\beta_{2_\mu})$     & -0.124 &  0.105 &  -0.295 &    0.047 &   35386 &    1.0 \\
$\Delta(\beta_{3_\mu})$     &  0.167 &  0.128 &  -0.045 &    0.371 &   50705 &    1.0 \\
$\Delta(σ)$      &  0.030 &  0.027 &  -0.014 &    0.074 &   16452 &    1.0 \\
\bottomrule
\end{tabular}
\label{exp2_parameter_estimates}
\end{threeparttable}
\end{center}
\end{table}

\end{document}